Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-12-22 Paper/ms.tex
Updated fake data results

\section{Demonstration on Simulated Data}\label{sec:fake_data}
To test our proposed methodology, we generated a simulated dataset for a
system with $N=5$ planets. The dataset was ``inspired by'' the $\nu$ Oph
dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
periods of 531 and 3180 days, whose semi-amplitudes are
322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. The other three planets have
much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
for the noise in the data was 5 m s$^{-1}$, so some of these low-amplitude
signals should be detectable. The simulated data is shown in
Figure~\ref{fig:fake_data}, along with the true radial velocity curve $m(t)$
that was used to generate the data.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data.eps}
\caption{A simulated dataset that ``resembles'' the $\nu$ Oph dataset, which
we used to test our methodology. The dominant signal is from two large
planets with periods periods of 531 and 3180 days and semi-amplitudes of
322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. There are also three
much smaller planets which contribute small additional effects to the data.
\label{fig:fake_data}}
\end{figure}

We ran the MCMC on the simulated dataset to obtain samples from the
posterior distribution. We obtained 590 samples.
The posterior distribution for $N$, the number of
planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
5, is also the most probable value according to the posterior distribution,
simply by chance. The posterior distribution also suggests that $N=4$ has
substantial probability.
The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
the posterior distribution for each period is identical, so we pooled the
samples for all periods. Defining the log-periods by
$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
The posterior distribution for the periods, shown in
Figure~\ref{fig:fake_data_periods}, shows that four of the true periods were
recovered, with probability close to 1. One period (with $\log_{10}P_i \approx 1.8$)
which was actually present was not ``detected'' because it had a very small
amplitude. We note that the posterior probability near this period should not be
precisely zero. There is also some
evidence for periods which did not actually exist, however the posterior
probabilities for these peaks are not close to 1.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
\caption{The posterior distribution for the number of planets $N$ given the
simulated dataset. The true number of planets was 5, which is also the most
probable value (this is a coincidence and will not occur in general).
\label{fig:fake_data_N}}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
\caption{The posterior distribution for the periods given the simulated
dataset. The solid lines are the true periods, four of which were detected
with high posterior probability.\label{fig:fake_data_periods}}
\end{figure}

The joint posterior distribution for the periods and the amplitudes of the
signals is shown in Figure~\ref{fig:fake_data_posterior} along with the
eccentricities. As with Figure~\ref{fig:fake_data_periods}, the samples for
all planets were combined. The true values are also plotted as circles. Clearly, the
reason the period of $\log_{10}(P_i) \approx 1.8$ was not ``detected'' was that
it had a very low amplitude of approximately 1 m s$^{-1}$ which is one fifth
of the noise level.

>>>>

\section{Demonstration on Simulated Data}\label{sec:fake_data}
To test our proposed methodology, we generated a simulated dataset for a
system with $N=7$ planets. The dataset was ``inspired by'' the $\nu$ Oph
dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
periods of 530 and 3120 days, whose semi-amplitudes are
291 m s$^{-1}$ and 181 m s$^{-1}$ respectively. The other five planets have
much lower semi-amplitudes, ranging from 4-30 ms$^{-1}$. The standard deviation
for the noise in the data was 5 m s$^{-1}$, so some of these low-amplitude
signals should be detectable. The simulated data is shown in
Figure~\ref{fig:fake_data}, along with the true radial velocity curve $m(t)$
that was used to generate the data.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data.eps}
\caption{A simulated dataset that ``resembles'' the $\nu$ Oph dataset, which
we used to test our methodology. The dominant signal is from two large
planets with periods periods of 530 and 3120 days and semi-amplitudes of
291 m s$^{-1}$ and 181 m s$^{-1}$ respectively. There are also five
much smaller planets which contribute small additional effects to the data.
\label{fig:fake_data}}
\end{figure}

We ran the MCMC on the simulated dataset to obtain samples from the
posterior distribution. We obtained 520 posterior samples.
The posterior distribution for $N$, the number of
planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
7, is not the most probable value, but it does have substantial probability.
The posterior distribution suggests that $N$ could be anywhere from 6 to 10.
The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
the posterior distribution for each period is identical, so we pooled the
samples for all periods. Defining the log-periods by
$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
The posterior distribution for the periods, shown in
Figure~\ref{fig:fake_data_periods}, shows that six of the true periods were
recovered, with probability close to 1. One period (with $\log_{10}P_i \approx 1.15$)
which was actually present was not ``detected'' because it had a very small
amplitude. We note that the posterior probability near this period should not be
precisely zero. There is also some
evidence for periods which did not actually exist, however the posterior
probabilities for these peaks are not close to 1.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
\caption{The posterior distribution for the number of planets $N$ given the
simulated dataset. The true number of planets was 7.
\label{fig:fake_data_N}}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
\caption{The posterior distribution for the periods given the simulated
dataset. The solid lines are the true periods, six of which were detected
with high posterior probability.\label{fig:fake_data_periods}}
\end{figure}

The joint posterior distribution for the periods and the amplitudes of the
signals is shown in Figure~\ref{fig:fake_data_posterior} along with the
eccentricities. As with Figure~\ref{fig:fake_data_periods}, the samples for
all planets were combined. The true values are also plotted as circles. Clearly, the
reason the period of $\log_{10}(P_i) \approx 1.15$ was not ``detected'' was that
it had a very low amplitude of approximately 4 m s$^{-1}$ which is below the
noise level.
